\LARGE \bf Approximately Orchestrated Road Traffic Automata\\ a General Traffic Simulation Framework

### 2012-02-19 papers/itsc_2012/root.tex
checking in my final comments on the first draft

@@

\commentpk{Needs 1 more para based on final application, and image on the top right, and introduction to the rest of the paper here once we finalize the sections}

@@

Another popular open source traffic simulator that can use OSM data is SUMO \cite{SUMO2011}, which shares the same goals as AORTA and is fairly similar at a high level. SUMO has been used to study the effect of automated transportation systems, route planning of individual vehicles, as well as dynamically adapting traffic signal policies to increase traffic efficiency. AORTA and SUMO differ in the manner in which intersections are handled. SUMO uses traditional traffic signals, while AORTA adopts a much more general \textit{reservation} system. A reservation tells a car when it is safe to take its desired path through the intersection. Reservations can be alotted to vehicles through more traditional means such as traffic signals or stop-sign based intersection based precedence, or through more efficient methods designed specificaly for autonomous intersection management \cite{JAIR08-dresner}. \commentpk{does this difference seem a bit pedantic?}

@@

\commentpk{I suspect there are still more papers that might be more relevant - we'll definitely have some when we write the application}

% In the realm of applications
% It is also necessary to analyze some other simulators that use GIS data for microscopic traffic simulations, this is because it may be possible to export OSM data so that it can be used by these simulators, though the authors are not aware of such a connection. Incident response has been studied by the following paper (though perhaps only a minor citation as it is an application and does not really compare against AORTA, but it is something we would like to study this as well. \cite{huang2007gis}

@@

\pix{architecture.eps}{An illustration of AORTA's components.}{scale=0.3}

The architecture and implementation notes of AORTA follow. AORTA is divided into
roughly three components, each somewhat independent of the others. The map model
handles everything from building graphs from OSM source to pathfinding. This
component is cleanly separated from the simulation engine, which adds the notion
of agents, structures on top of map components that detect collisions, and the
machinery to implement AORTA's model of car dynamics. If a future experiment
demands macrosimulation or cellular automata method, no changes to the map
infrastructure are needed, due to the software engineering. Finally, there is a
UI using Java Swing, which is (out of necessity) tied to both the map and
simulation components. The software has a documented API, is open-source, and
contains all of the code for the experiments described in Section 6 -- it is
designed for extensibility.

@@

There are three standalone applications in AORTA. The map builder takes an OSM
file as input and produces an AORTA graph as output. There is a UI, using Swing,
to interactively explore the map and run simulations. Specific traffic patterns
can quickly be simulated by selecting a source and destination group of roads in
the UI. Finally, there is a headless mode that runs some specific experiment
without the overhead of visualization. The UI and headless mode both take a
graph as input, meaning the map construction step must be run once per OSM
input, but the same graph can be reused indefinitely.

@@

@@

Finally, coordinates are normalized from GPS (longitude, latitude) to a
Cartesian coordinate system with no negative coordinates and the Y coordinate
increasing upwards. This is only done for convenience of manipulation in the UI.
Any coordinate is translated back to its original GPS form when used for
computing traversable length.

@@

AORTA's structure demands a traditional definition of edge linking exactly two
vertices, rather than OSM's ways going through many common nodes. Pass 2 walks
along each way's coordinates, dividing it into multiple undirected edges.

@@

@@

Next, turns are constructed at each vertex between the incoming and outgoing
edges. This is completely heuristic based: we classify a potential turn as a
left, right, or straight turn based solely on angles between edges and if the
edge's original OSM ID matches. U-turns are constructed at dead-ends. If several
lanes all cross into fewer lanes, merging is forced for the rightmost lanes.
These patterns are attempts at generalizations; this process results in many
intersections having the incorrect turns. However, without a more detailed
source, this is the best compromise. At the time of writing, even sources like
Google Maps lack this information. (TODO how do you cite a lack of something?)
Finally, the geometry of a turn is simplified to a single line between the ends
of the edges it connects.  In the future, a more realistic curve could be drawn.

@@

AORTA has a low-granularity model of turn conflict. If two vehicles at any point
along two different turns could ever come into contact, then the turns are
always in conflict, no matter where the agent specifically is along that turn.
The alternative of tiling intersections (like \cite{JAIR08-dresner} does) would
provide more granularity, but it adds computational complexity and would be
impractical for many of the complex intersection geometries inferred from OSM.

@@

Because turns are constructed imperfectly, most large graphs have edges that are
not reachable from some portions of the graph. Simulation and pathfinding would
suffer greatly if this remained the case, so pass 3 uses Tarjan's linear-time
algorithm for locating strongly connected components to remove the components of
the directed graph with the fewest members. (There is also a debugging switch to
leave these components in and highlight them in the UI.) The algorithm operates
at the granularity of lanes. If several lanes of many are disconnected within
one road, then the bad lanes are spliced out. This results in visually
unappealing geometric gaps, but if the entire road were removed, then Tarjan's
algorithm would have to be repeated to determine what edges the good lanes would
then disconnect. The end result is a connected digraph, ideal for corner
case-free pathfinding and simulation.

@@

\commentd{this maybe doesn't belong in this paper}

@@

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Discrete-time simulation is used, meaning an agent accelerates at a fixed rate
for the duration of a ``tick'' (also referred to as ``step'') to achieve a new
position and velocity. The tick duration $dt$ is configurable and fixed for the
simulation. Agents know this value and use it to reason about collision
avoidance and satisfying other constraints, so setting this value to the
average ``reaction time'' of a human driver may be appropriate. Each tick, the
simulation performs the following steps:

@@

@@

There are two structures to detect collisions. All traversables (edges and the
straight line-approximation of turns) maintain a \emph{queue} of occupying
agents, ordering them so that the agent farthest along is at the head of the
queue. A naive method of collision detection would detect an agent passing an
agent on the same queue, or entering a new queue not at the end, immediately
after an agent moves. However, agents take their steps sequentially and in an
effectively random order. Ideally they could be reordered so that agents in the
front of a queue always move first. Since the first agent may cross into the
next queue, that introduces a depdendency between traversables. A topological
sort of a directed graph containing circuits is not possible though, and since
we guarantee connectedness, there are necessarily circuits in maps.  Instead,
collisions are checked after every agent has stepped forward in time, treating
each time-step as atomic. Queues remember the original order of agents before a
step, and afterwards can verify that the order is still legal, with new agents
on the end and existing agents in the same order.

@@

\commentd{include some form of piyush's proof about accel to end?}

@@

When the lookahead analysis is considering some future traversable, it only has
to consider the agent closest to the start of it in the present. An edge may
only be entered from a turn, so the intersection policy has the burden of
ensuring no agent performs a turn that leads to the same edge simultaneously.
Following an agent safely means maintaining a configurable following distance
behind the agent. Worst-case analysis is again used: the most dangerous move the
agent in front could perform is decelerating as fast as possible the next tick,
and the worst-case traveling distance for the agent following is already known
from lookahead (speeding up as much as possible).

@@

\commentd{give some informal or formal (with numbers) idea of scale / how many
agents we can simulate in different map sizes}

>>>>

% \long\def\commentp#1{}
% \long\def\commentpk#1{}
% \long\def\commentm#1{}
% \long\def\commentd#1{}

@@

\commentpk{Needs 1 more para based on final application}

@@

Another popular open source traffic simulator that can use OSM data is SUMO
\cite{SUMO2011}, which shares the same goals as AORTA and is fairly similar at
a high level. SUMO has been used to study the effect of automated
transportation systems, route planning of individual vehicles, as well as
dynamically adapting traffic signal policies to increase traffic efficiency.
AORTA and SUMO differ in the manner in which intersections are handled. SUMO
uses traditional traffic signals, while AORTA adopts a much more general
\textit{reservation} system. A reservation tells a car when it is safe to take
its desired path through the intersection. Reservations can be allotted to
vehicles through more traditional means such as traffic signals or stop-sign
based intersection based precedence, or through more efficient methods designed
specifically for autonomous intersection management \cite{JAIR08-dresner}.
\commentpk{does the above difference seem a bit pedantic?}

@@

@@

\pix{architecture.pdf}{An illustration of AORTA's components.}{scale=0.3}

\commentpk {This section is needlessly verbose. I recommend reducing the text down to 1 small para.}

The architecture and implementation notes of AORTA follow. AORTA is divided
into roughly three components, each somewhat independent of the others. The map
model handles everything from building graphs from OSM source to pathfinding.
This component is cleanly separated from the simulation engine, which adds the
notion of agents, structures on top of map components that detect collisions,
and the machinery to implement AORTA's model of car dynamics. If a future
experiment demands macro-simulation or cellular automata method, no changes to
the map infrastructure are needed, due to the software engineering
\commentpk{just saying software engineering seems a bit odd} . Finally, there
is a UI using Java Swing, which is (out of necessity) tied to both the map and
simulation components. The software has a documented API, is open-source, and
contains all of the code for the experiments described in Section 6 -- it is
designed for extensibility. \commentpk{Insert a footnote to the code here instead od using an Appendix}

@@

\commentpk{the next para could be merged with the first one} There are three
standalone applications in AORTA. The map builder takes an OSM file as input
and produces an AORTA graph as output. There is a UI, using Swing, to
interactively explore the map and run simulations. Specific traffic patterns
can quickly be simulated by selecting a source and destination group of roads
in the UI \commentpk{I would emphasize the demand generation a bit more}.
Finally, there is a headless mode that runs some specific experiment without
the overhead of visualization. The UI and headless mode both take a graph as
input, meaning the map construction step must be run once per OSM input, but
the same graph can be reused indefinitely.

@@

\commentpk{Since OSM data has been used in the past in a couple of approaches,
I would recommend re-writing this entire section down to a 2-3 paragraph
sub-section. This is sort of the infrastructural hard work that gets swept
under the rug, but you can write about it in future papers (journal articles
etc). Remember to explicitly save a copy of the text before editing. I have
written a few comments that can potentially be used in the future}

@@

Finally, coordinates are normalized from GPS (longitude, latitude) to a
Cartesian coordinate system with no negative coordinates and the Y coordinate
increasing upwards. This is only done for convenience of manipulation in the UI.
Any coordinate is translated back to its original GPS form when used for
computing traversable length.\commentpk{Is your cartesian coordinate system arbitrary? If so, I would highly recommend using the UTM system (google wiki UTM). We already have the code to convert to it from lat/long coordinates.}

@@

AORTA's structure demands a traditional definition of edge linking exactly two
vertices, rather than OSM's ways going through many common nodes \commentpk{I am not sure what you mean by the first statement}. Pass 2 walks
along each way's coordinates, dividing it into multiple undirected edges.

@@

\commentpk{see if you can remove the passive voice. Maybe name the utility \textit{map\_builder} and use that instead}

@@

Next, turns are constructed at each vertex between the incoming and outgoing
edges. This is completely heuristic based: we \commentpk{don't use we} classify a potential turn as a
left, right, or straight turn based solely on angles between edges and if the
edge's original OSM ID matches. U-turns are constructed at dead-ends. If several
lanes all cross into fewer lanes, merging is forced for the rightmost lanes.
These patterns are attempts at generalizations; this process results in many
intersections having the incorrect turns. However, without a more detailed
source, this is the best compromise. At the time of writing, even sources like
Google Maps lack this information. (TODO how do you cite a lack of something?)
Finally, the geometry of a turn is simplified to a single line between the ends
of the edges it connects.  In the future, a more realistic curve could be drawn.

@@

\commentpk{The following paragraph should not be here. It should be in intersection behaviors}, AORTA has a low-granularity model of turn conflict. If two vehicles at any point
along two different turns could ever come into contact, then the turns are
always in conflict, no matter where the agent specifically is along that turn.
The alternative of tiling intersections (like \cite{JAIR08-dresner} \commentpk{I made this same mistake a month ago, dont use citations as a noun} does) would
provide more granularity, but it adds computational complexity and would be
impractical for many of the complex intersection geometries inferred from OSM.

@@

Because turns are constructed imperfectly, most large graphs have edges that are
not reachable from some portions of the graph. Simulation and pathfinding would
suffer greatly if this remained the case, so pass 3 uses Tarjan's linear-time
algorithm for locating strongly connected components to remove the components of
the directed graph with the fewest members \commentpk{missing citation?}. (There is also a debugging switch to
leave these components in and highlight them in the UI.) The algorithm operates
at the granularity of lanes. If several lanes of many are disconnected within
one road, then the bad lanes are spliced out. This results in visually
unappealing geometric gaps, but if the entire road were removed, then Tarjan's
algorithm would have to be repeated to determine what edges the good lanes would
then disconnect. The end result is a connected digraph, ideal for corner
case-free pathfinding and simulation.

@@

\commentd{this maybe doesn't belong in this paper} \commentpk{yes. You can state that OSM data is not perfect, but I would definitely remove this for now}

@@

\commentpk{This needs to be removed as well, it should be in future work}

@@

AORTA uses microscopic simulation, modeling individual drivers as agents.
Discrete-time simulation is used, meaning an agent accelerates at a fixed rate
for the duration of a ``tick'' (also referred to as ``step'') to achieve a new
position and velocity. The tick duration $dt$ is configurable and fixed
\commentpk{configurable \textit{and} fixed is confusing} for the simulation.
Agents know this value and use it to reason about collision avoidance and
satisfying other constraints, so setting this value to the average ``reaction
time'' of a human driver may be appropriate. Each tick, the simulation performs
the following steps:

@@

\commentpk{I wonder if demand data generation in the ui should go here.}

@@

There are two structures to detect collisions. All traversables (edges and the
straight line-approximation of turns) maintain a \emph{queue} of occupying
agents, ordering them so that the agent farthest along is at the head of the
queue. A naive method of collision detection would detect an agent passing an
agent on the same queue, or entering a new queue not at the end, immediately
after an agent moves. However, agents take their steps sequentially and in an
effectively random order. Ideally they could be reordered so that agents in the
front of a queue always move first. Since the first agent may cross into the
next queue, that introduces a dependency between traversables. A topological
sort of a directed graph containing circuits is not possible though, and since
we guarantee connectedness, there are necessarily circuits in maps.  Instead,
collisions are checked after every agent has stepped forward in time, treating
each time-step as atomic. Queues remember the original order of agents before a
step, and afterwards can verify that the order is still legal, with new agents
on the end and existing agents in the same order.

@@

\commentd{include some form of piyush's proof about accel to end?} \commentpk{I would say it is unnecessary}

@@

When the lookahead analysis is considering some future traversable, it only has
to consider the agent closest to the start of it in the present. An edge may
only be entered from a turn, so the intersection policy has the burden of
ensuring no agent performs a turn that leads to the same edge simultaneously.
Following an agent safely means maintaining a configurable following distance
behind the agent \commentpk{not sure what you mean by configurable here}. Worst-case analysis is again used: the most dangerous move the
agent in front could perform is decelerating as fast as possible the next tick,
and the worst-case traveling distance for the agent following is already known
from lookahead (speeding up as much as possible).

@@

\commentd{give some informal or formal (with numbers) idea of scale / how many
agents we can simulate in different map sizes} \commentpk{I would say that the the last point is fairly important}
